\bf Estimating the number of clusters using cross-validation

### 2016-06-18 cvclust.tex
Correct a few typos

A clustering procedure segments a collection of items into smaller groups,
with the property that items in the same group are more similar to each other
than items in different groups \citep{hartigan1975clustering}.  Such
procedures are used in two main applicaitons: (a) exploratory analysis, where clusters
reveal homogeneous sub-groups within a large sample; (b) data
reduction, where high-dimensional item attribute vectors get reduced discrete
cluster labels \citep{jain1999data}.

@@

The problem of choosing $k$ has been well-studied, and dozens of methods have
been proposed~\citep{chiang2010intelligent,fujita2014non}. The main difficulty
in choosing $k$ is that clustering is fundamentally an ``unsupervised''
learning problem, meaning that there is no obvious way to use ``prediction
ability'' to drive the model selection \citep{hastie2009elements}.  Most
existing methods for choosing $k$ instead rely on on explicit or implicit
assumptions about the data distribution, including it shape, scale, and
correlation structure.

Several authors advocate choosing $k$ by performing a sequence of hypothesis
tests will null and alternative hypotheses of the form $H_0 : k = k_0$ and
$H_1: k > k_0$, starting with $k_0 = 1$ and proceeding sequentially with
higher values of $k_0$ until a test fails to reject $H_0$. The gap statistic
method typifies this class of methods, with a test statistic that measures the
within-cluster dispersion relative to what is expected under a reference
distribution \citep{tibshirani2001estimating}.
%Similar methods include those developed by \citet{TODO}.

@@

This setting investigates performance in the presence of heavy-tailed data.
When the degrees of freedom decreases, the tail becomes more flat and the
Gaussian assumption becomes more appropriate.  For most methods, their
performances are relatively stable until the tail gets very heavy. In the case
where there are $2$ degrees of freedom, the Gap and Jump methods' performances
deteriorate considerably relative to the Gabriel method.

@@

The first data set is congressional voting data consisting of voting records
of the second session of the $98$th United States Congress,
\citep{schlimmer1987concept}.  This data set includes votes for legislators on
the $P = 16$ key votes.  For each vote, each legislator either votes
positively (``yea'') negatively (``nay''). We removed legislators with missing
votes.  This results in $N = 232$ remaining records, with $124$ democrat and
$108$ republican. There are $k = 2$ clusters of legislators, corresponding to
political party.

>>>>

A clustering procedure segments a collection of items into smaller groups,
with the property that items in the same group are more similar to each other
than items in different groups \citep{hartigan1975clustering}.  Such
procedures are used in two main applicaitons: (a) exploratory analysis, where clusters
reveal homogeneous sub-groups within a large sample; (b) data
reduction, where high-dimensional item attribute vectors get reduced to discrete
cluster labels \citep{jain1999data}.

@@

The problem of choosing $k$ has been well-studied, and dozens of methods have
been proposed~\citep{chiang2010intelligent,fujita2014non}. The main difficulty
in choosing $k$ is that clustering is fundamentally an ``unsupervised''
learning problem, meaning that there is no obvious way to use ``prediction
ability'' to drive the model selection \citep{hastie2009elements}.  Most
existing methods for choosing $k$ instead rely on explicit or implicit
assumptions about the data distribution, including it shape, scale, and
correlation structure.

Several authors advocate choosing $k$ by performing a sequence of hypothesis
tests with null and alternative hypotheses of the form $H_0 : k = k_0$ and
$H_1: k > k_0$, starting with $k_0 = 1$ and proceeding sequentially with
higher values of $k_0$ until a test fails to reject $H_0$. The gap statistic
method typifies this class of methods, with a test statistic that measures the
within-cluster dispersion relative to what is expected under a reference
distribution \citep{tibshirani2001estimating}.
%Similar methods include those developed by \citet{TODO}.

@@

This setting investigates performance in the presence of heavy-tailed data.
When the degrees of freedom decreases, the tail becomes more flat and the
Gaussian assumption becomes more inappropriate.  For most methods, their
performances are relatively stable until the tail gets very heavy. In the case
where there are $2$ degrees of freedom, the Gap and Jump methods' performances
deteriorate considerably relative to the Gabriel method.

@@

The first data set is congressional voting data consisting of voting records
of the second session of the $98$th United States Congress,
\citep{schlimmer1987concept}.  This data set includes votes for legislators on
the $P = 16$ key votes.  For each vote, each legislator either votes
positively (``yea'') or negatively (``nay''). We removed legislators with missing
votes.  This results in $N = 232$ remaining records, with $124$ democrat and
$108$ republican. There are $k = 2$ clusters of legislators, corresponding to
political party.
